Efficient Computation of Extended Surface Sources

### 2023-01-25 repro/papers/GEM22/submitR2/response.tex

COMMENTS FOR THE AUTHOR:

Dear Author, 
Thanks for this revised version of your  study. Both reviewers acknowledge the significant improvement of this new version. They still point some typos, and comments/questions I think would be beneficial to consider in a further revised version of your study before we accept it for publication. See these comments appended below. 
Thanks a lot for your work and this interesting study, 
Best regards,
Ludovic Métivier.

=============================================================================================================================================================================================================================

Reviewer #1: The manuscript has been considerably revised. The revised version is much better. The early introduction of the examples and the overview of the time reversal method significantly improve the readability. I would like to recommend the manuscript for publication. 
I have, however, two questions and I have detected typos (I am afraid that there are many more).

Comment 1. what are the blue and red rays in Figure 3 ?
the blue rays start from the source, the red rays start from a point located just below the source. As a result, some red and blue rays are very close to each other.

Response: I have created a clearer ray diagram overlain on the bulk modulus color scale plot (Figure 3) and added another figure and some discussion to explain the ray geometry. The red and blue rays start at the same source point and are simply different parts of the same point source ray field. The diving wave example is similar to that depicted in Figure 4.20a in Sheriff and Geldart, {\em Exploration Seismology}. It features a homogeneous upper layer, a middle layer in which the velocity increases rapidly, and a homogeneous lower layer, with smooth transitions between the layers. As Sheriff and Geldart say, ``A velocity gradient in a layer that is substantially higher than the velocity gradient above the bed, as in fig. 4.20a, causes a very sharp increase in ray path curvature and a folding back of the time-distance curve. Such a triplication of branches of the time-distance curve (fig. 4.20c) is usually difficult to see because the later branches become lost in the later cycles of the earlier arrivals.'' To show clearly how this triplication happens, I've included a plot of arrival times in my introduction of this example (new Figure 4b). The part of the arrival time curve in red corresponds to the those red rays in the ray plot (Figure 3) that refract back to the receiver surface, and similarly for the blue rays. Note that the red branch is close enough to the direct arrival time that its chief effect is an apparent wavelet change in the direct arrival. Both red and blue branches are really diving wave arrival times, but only the waveform arrival corresponding to the blue branch can be isolated by a simple mute.

Comment: on page 23, I do not understand the following sentence:
The function a_0 is a frequency-independent envelope, equal to 1 where h is non-zero.

Response: This section skipped over a number of steps, which I have included in the revision. I believe that I have explained this idea.

typos:

page 2
certainluy
->
certainly

page 5
veloity
->
velocity

propblem
->
problem

tahnks 
->
thanks to

is proportion to 
->
is proportional to

page 6

slowiy
->
slowly

page 8
increasting
->
increasing

page 10

to generate a data traces

page 13

of sigma 
-> 
of $\Sigma$

page 16
representaion 
->
representation

page 21
too many s in Psshat{p}

page 22
Speclializing
->
Specializing

last equation: confusion between s and t
partial v partial t
the use of s is very confusing (it is the symbol of surface !)

page 23
first equation: confusion between s and t

SInce 
->
Since

check the sentence:
is asymptotic the pressure field of to an incoming solution

last equation: confusion between s and t

page 24
parenthesis in lambda(...

page 26
parenthesis in (tilde{p},tilde{v}

negligbly -> negligibly

defintiion -> definition

page 29
as it need  
->
as it needs

to ne discussed
->
to be discussed

page 31
parenthesis in (equations 28,29

page 33
typo in: and equation 48, .

page 34
computationa
->
computational

page 36
ismeasured 
->
is measured

page 37
parenthesis in (Figure 15b

page 39
Madacascar
->
Madagascar

=============================================================================================================================================================================================================================

Reviewer #2: Overall the paper is much improved and I now found it relatively easy to follow and understand.  I particularly liked the 'steps' introduced as a way to summarize what you've done so far and to give the reader further help in actually implementing these ideas.  (I also liked the roadmap section at the beginning.)  I also found the new figures much easier to compare and understand.

There are a few areas where I think things could still be somewhat clearer, which are:

Comment 1. In the introduction p 3 line 44: Perhaps say in a few words why you isolate the diving rays.  Also later when you show your ray picture (Figure 3) with the diving waves, first I think I see red rays beside the blue ones, and second just out of curiosity, what is your definition of diving wave, because I see things that turn back to the surface that you do not deem diving rays.

Response: You are of course correct that the ``things that turn back'' are another diving wave branch. I've considerably revised the figure, its caption, and the discussion in the text, intending to answer all of these concerns.

Comment 2. You relate what you are doing to photo-acoustic tomography, which is neat and certainly was the motivation.  As I read the paper though, I can't help thinking that what you are doing is a lot like running RTM back to the source.  I expect that other readers may also wonder what is really fundamentally different between RTM and this approach, so I think that adding something to make this comparison explicit would be helpful.

Response: this is a very good question. I have added a paragraph describing the role of the D-to-N map in ``true amplitude RTM'', aka approximate Born inversion. In fact it's quite similar to the role described in this paper for the inverse surface source problem. In both cases the core idea is to backpropagate the actual data field. In the present case, that is a transmitted field; in true-amplitude migration, it's the scatterered field. Reconstructing these from the data is the key to inversion in both cases, and the D-to-N map is the key to reconstruction.

Comment 3. In figure 2 and the lens example in general, why is the source you used not on the source surface?  Perhaps I have misunderstood the meaning of the source surface, and that could be clarified.  I guess that the source should be below that surface so the rays come into it from below and then leave it 'from above' and enter the receiver surface from below?   I am sure that one could phrase that much better than I have here, but as I was reading it was odd to see the source below what's called the source surface.

Response: I could have placed the source on the source surface. I created a parallel sequence of experiments for the lens example using that configuration. However a point source cannot be reconstructed, at least not with the simple transmission geometry used in this paper, since a lot of energy propagates away ``to China'' and is not available in the inversion. You can reconstruct a source that generates the data, but it isn't a point source - if the velocity is right, it is more or less a dip-filtered version of the point source. Instead, I chose to somewhat artificially create a source that can actually be reconstructed fairly accurately in this example: the rays carrying significant energy actually make it to the receiver surface within the time limit.

The diving wave example uses a point source on the source surface. You will notice that the reconstructed source (Figure 9b) appears to be a dip-filtered point source, regarded as a function on the source surface.

Comment 4, In Figure 3, there are some rays that look weird to me, at around (8000,1000) and (12000,1000) This is just a gradient so I am expecting more-or-less circular wavefronts.  I don't know that this really requires a comment or anything, but maybe a mention as to what is going on there.

Response: Figure 3 and surrounding discussion needed some tuning up. Note that the gradient-to-constant transition is smoothed, and the gradient is large enough to generate a triplication. Having homogeneous layers top and bottom really messes things up. Sheriff and Geldart, {\em Exploration Geophysics}, discussion of fig. 4.20a, have a nice discussion of this behaviour. There are segments of circles in there, but overall these rays are not close to circles. I think the description now in the text, and the inclusion of the arrival time plot, will make things clear enough.

Comment 5. p 11 line 32, what do you mean by 'must be localized' in which variables?

Response: The PAT construction does not apply to the surface source inversion problem, since there is no bounding surface on all of which the pressure field is measured/known. I get around this by localizing in space-time. For short-time propagation, the signal does not propagate far from the source surface, so an artificial bounding surface can be added on which the field is automatically zero. So the field can be viewed as the solution of a local pressure boundary value problem, for a little while. Then it can be converted to a local solution of a problem with surface source, as described. Since the field vanishes near the artificial boundary, you can simply get rid of it, and if the pressure field on the boundary is only non-zero for a short time, that's it - just let it propagate in space time. Finally, any pressure distribution on the source surface can be broken up in to a sum of terms, each only non-zero for a short interval. Apply the short-time-window construction to each summand, and add them up.

Comment 6. Figure 6b has quite a different source than that shown in Figure 2b.  I think that when you talk about Figure 6 it would be good to re-iterate why you are getting a different source here (I think velocity vs pressure, but remembering that many people only look at the pictures this will be confusing).

Response: I think this is actually explained in the caption of Figure 7 (formerly Figure 6), also in the text discussion. This is the result of inversion {\em in the wrong velocity}. The data (Figure 2a or 2c) is inverted here in a homogeneous model - so the source does not look at all like that (2b) used to generate the data in the lens model.

Typos: I am {\em very} grateful to both referees for taking the trouble to list many of the embarassingly extensive typos. I believe that I have fixed all of these, and some others which we won't mention. I only hope I have not created yet more...

A very-incomplete typo list (I do suggest getting out your grammatical fine-toothed comb and checking for more of these):

-Spelling of constraint in the abstract
-p 6 line 23 "is not be well-defined"
-p 13 line 24 there is a sigma without a \
-p 16 line 59 spelling of representation
-P 21 line 40 The sentence beginning "Differentiate" is missing some important part-of-speech for which I should have a name but don't, perhaps a subject?
-p26 line 28 missing )
-p26 line 48 sentence fragment beginning "To illustrate"
-p26 line 62 missing space in is measured

### 2023-01-25 repro/papers/GEM22/submitR2/paper.tex
pre-sub

\documentclass[georeport,12pt]{geophysics}

@@

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{alg}{Algorithm}
\newtheorem{remark}{Remark}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{definition}{Defenition}

@@

\maketitle
\begin{abstract}
Source extension is a reformulation of inverse problems in wave propagation, that at least in some cases leads to computationally tractable iterative solution methods. The core subproblem in all source extension methods is the solution of a linear inverse problem for a source (right hand side in a system of wave equations) through minimization of data error in the least squares sense with soft imposition of physical constraints on the source via an additive quadratic penalty. A variant of the time reversal method from photoacoustic tomography provides an approximate solution that can be used to precondition Krylov space iteration for rapid convergence to the solution of this subproblem. An acoustic 2D example for sources supported on a surface, with a soft contraint enforcing point support, illustrates the effectiveness of this preconditioner.
\end{abstract}

@@

The modeling operator $PL[{\bf c}]^{-1}$ and the penalty operator $A$ defined in \ref{eqn:penop} are linear, so the source
subproblem is a linear least squares problem. Under some additional
assumptions to be described below, I shall show how to construct an
accurate approximate solution operator for problem
\ref{eqn:esis}. This approximate solution operator may be used to
accelerate (``precondition'') Krylov space methods for the solution of the surface source
subproblem \ref{eqn:esis}. I will fully describe a preconditioner for a special
case of the source subproblem \ref{eqn:esis}, in which ${\bf u}$ is an
acoustic field, $L[{\bf c}]$ is the wave operator of linear
acoustodynamics, and the source and receiver surfaces are horizontal
planes. Numerical examples in this setting suggest the
effectiveness of this acceleration.

@@

A simplified mathematical
translation of this medical imaging task is to infer the initial
excess pressure distribution over a fluid-containing region at time
$t=0$ from measurements of the pressure on a surface enclosing the
fluid over a time interval $0 \le t \le t_{\rm max}$. The time reversal method presumes that the
pressure field has returned to equilibrium (zero excess pressure), or
close to it, in the fluid-containing region at the
final time $t=t_{\rm max}$ Then the field can be (at least
approximately) viewed as the backwards-in-time initial boundary value
problem with zero final conditions at $t=t_{\rm max}$, and boundary
values given by the measurements. Evolving the field backwards in time
backwards to $t=0$ thus solves the problem. Except in special
circumstances, the pressure field never actually vanishes at finite
time, so the solution is approximate.

@@

First, reverse-time propagation can be localized via ray theory,
within high-frequency asymptotic approximation. This step requires
some assumptions about the ray fields: all rays responsible for
significant energy in the receiver data must arrive from the source
surface, and must cross the source and receiver surfaces
transversally, that is, making non-zero angles. Also, the rays must
approach source and receiver surfaces from one side or the other
(``the inside''), so that locally the surfaces can be treated as
boundaries of propagation domains. Specifically, rays must emanate
from the source surface towards the inside (``incoming'' rays), and
arrive at the receiver surface from the inside (``outgoing'' rays).
With these assumptions, reverse-time propagation of the receiver data
closely approximates the acoustic fields near the source surface. Next
observe that for a causal field with a surface pressure source (with a
continuous pressure field across the surface), the source is
proportional to the jump in the (particle) velocity field. Moreover,
to leading order in frequency, the velocity field switches sign at the
surface, so the jump is just twice the sampled value on the
surface. Thus reverse-time propagation and reading off the velocity
field on the source surface yields a source that reproduces the
pressure data, within an asymptotically small error.
%Preconditioning amounts to a change of
%norm in the domain and range spaces of the modeling operator. In this
%case, the modfied norms are weighted $L^2$, and the weight operators
%map pressure to corresponding surface source on the source and
%receiver planes.

This is the essence of the approximate solution of the problem
\ref{eqn:esis} for penalty weight $\alpha=0$. To see how Krylov
iteration might be accelerated, note that time-reverse propagation of
the receiver pressure values may be represented by a time-reverse
source propagation, with the source constructed from the velocity
field at the receiver surface. Moreover, the reverse-time
source-to-pressure propagation is the transpose (adjoint) of the
forward-time source-to-pressure propagation - this is a very simple
version of the adjoint state construction \cite[]{Plessix:06}. So the
approximate inverse may be described as: conversion of pressure to source at the
receiver surface, followed by the adjoint of the modeling operator,
followed by the conversion of pressure to source at the source
surface. This sequence precisely describes the adjoint of the modeling
operator with respect to weighted norms in the spaces of source and
receiver data, with the weight operators being the
pressure-to-source operators. That this construction results in an
approximate inverse means that the modeling operator is approximately
unitary with respect to these weighted norms, which in turn implies
that a properly constructed preconditioned conjugate gradient algorithm should converge
rapidly.

FInally, the $\alpha=0$ case is not sufficient: the penalty operator
$A$ is an essential component of the nonlinear inverse propblem
\ref{eqn:esi}. As it turns out, $A$ commutes with the other operators
involved - approximately, but that is enough. Therefore the effect of
$A$ can be compensated with an easily-computed factor, whence the
preconditioned acceleration extended to the case $\alpha > 0$.

This pressure-to-source map is closely related to the ``hyperbolic
Dirichlet-to-Neumann'' operator that plays a prominent role in
photoacoustic tomography and other wave inverse problems
\cite[]{Rachele:00,StefUhl:05}. \cite{HouSymes:EAGE16} demonstrated a
very similar preconditioner for Least Squares Migration, also for its
subsurface offset extension \cite[]{HouSymes:16}, motivated by
\cite{tenKroode:12}. These constructions also involve the
Dirichlet-to-Neumann operator. This concept also turns up in hidden
form in the work of Yu Zhang and collaborators on true amplitude
migration
\cite[]{YuZhang:14,TangXuZhang:13,XuWang:2012,XuZhangTang:11,Zhang:SEG09}.

%The obvious computation of the pressure-to-source map - prescribe the
%pressure, solve the wave equation with this boundary condition, read
%off the equivalent source - suffers from intrinsic numerical
%inaccuracy. I suggest an alternative computationally feasible
%approach, via economical short-distance wave propagation. Since the
%map is symmetric only in an approximate, asymptotic sense, it must be
%symmetrized for use as a Krylov preconditioner. I describe a
%symmetrization procedure that requires no further wave computations
%beyond those necessary to compute the action of the operator itself.

The discussion in this paper is formal and incomplete, in the sense
that some important mathematical underpinnings are only
sketched. I will treat the modeling operator $PL[{\bf c}]^{-1}$ as if
it mapped square integrable surface sources to square integrable sampled
data. This is not true in full generality: while the surface source
problem has distribution solutions, they are not generally square
integrable (finite acoustic field energy). Even if the solutions have
finite energy, they do not in general have well-defined
restrictions to lower-dimensional sets. In other words, the action of
the sampling
operator $P$ on the receiver surface is not well-defined for arbitrary
finite-energy acoustic fields. Thus the modeling operator envisioned above
is not be well-defined, strictly speaking.

This phenomenon is related to the ill-posedness of wave equations as
evolution equations in spatial variables, an observation attributed to
Hadamard (see \cite{CourHil:62}, Chapter 6, section 17). A number of
authors have described precise forms of the ray conditions described
above, and shown how these conditions lead to the desired behaviour of
the modeling operator, that is, mapping of finite energy sources to
finite energy data, or equivalent properties
\cite{Payn:75,Symes:83,Lasi:86,LasLionsTrig:86,Lasi:87, BaoSy:91b}.
Elaboration of these mathematical details is beyond the scope of this
paper, which aims instead to explore the algorithmic consequences of
the mathematical structure implied by the nongrazing hypothesis.

@@

Most of what follows is valid for any space dimension $n >0$. I will
describe the theory for $n=3$, write $\bx=(x,y,z)^T$ for the spatial
coordinate vector, and similarly for the 
particle velocity $\bv = (v_x,v_y,v_z)$. For computational convenience, the
examples are two-dimensional.

@@

For
acoustic modeling, surface sources are combinations of constitutive law
defects and loads normal to the surface, localized on $\Sigma_s$. That
is, right-hand sides in the system $L\bu=\bff$ take the form
$\bff(\bx,t) = (h_s\delta_{\Sigma_s},
f_s{\bf n}_s\delta_{\Sigma_s})^T$ for scalar defect $h_s$ and normal
load $f_s$. With the choice $L$ given in
\ref{eqn:awemat}, the causal system $L\bu=\bff$
takes the form
\begin{eqnarray}
\label{eqn:awepm}
\frac{1}{\kappa}\frac{\partial p}{\partial t} & = & - \nabla \cdot \bv +
h_s \delta_{\Sigma_s}, \nonumber \\
\rho\frac{\partial \bv}{\partial t} & = & - \nabla p +
                                                f_s{\bf n}_s \delta_{\Sigma_s},\nonumber \\
p & =& 0 \mbox{ for }  t \ll 0,\nonumber\\ 
\bv & = & 0 \mbox{ for } t \ll 0.
\end{eqnarray}

@@

Extended forward modeling consists in solving \ref{eqn:awepm} and
sampling the solution components at a receiver surface $\Sigma_r$ with
(``outward'') unit normal field ${\bf n}_r$.
$P_s$ and $P_r$ are the sampling operators on $\Sigma_s$ and
$\Sigma_r$ respectively. In
practice, sampling necessarily occurs at a discrete array of points
(trace locations), and over a zone of finite extent. In this
theoretical discussion, I will neglect the finite sample rate, and
regard the data, for example $P_rp$, as continuously sampled. The
output samples are necessarily muted, that is, non-zero only over a space-time domain
of finite extent. This mute, and any tapering applied to the data
traces, are regarded as part of the sampling operators $P_s,P_r$.

@@

Denote by $\Pi_i, i=0,1$ the projection on the first,respectively
second, component of a vector in $\bR^2$. The forward modeling
operator from pressure source to pressure trace is
\begin{equation}
  \label{eqn:sdef}
  S = \Pi_0 {\cal S}\Pi_0^T 
\end{equation}
%and the forward modeling operator from velocity source (normal force)
%to velocity trace is
%\begin{equation}
%  \label{eqn:vdef}
%  V_{z_s,z_r} = \Pi_1 {\cal S}_{z_s,z_r} \Pi_1^T 
%\end{equation}

@@

The two examples mentioned in the introduction illustrate the setting
just described. The first example (Figure \ref{fig:olensrays0}) embeds
an acoustic (low-velocity) lens between a source surface
$\Sigma_s= \{z = z_s=3000 \mbox{ m}\}$ (orange horizontal line) m and
a receiver surface $\Sigma_r= \{z=z_r=1000 \mbox{ m}\}$ (yellow
horizontal line).  The data to be used in this example results from a
point source at $(3500, 3500)$ m (orange star). I have overlain
the rays connecting this source point with the data portion of the
receiver surface: evidently all of the rays involved cross the source
and receiver surfaces transversally. Also, regarding ``inwards'' as
being ``upwards'' at the source surface, ``downwards'' at the receiver
surface, it is clear that the data is incoming at the source surface,
outgoing at the receiver surface.

\begin{figure}
  \includegraphics[width=\textwidth]{olensrays0.pdf}
  \caption{Bulk modulus with acoustic lens. Color scale unit is
    GPa. Orange horizontal line is source surface at depth $z_s= 3000$
    m, yellow horizontal
    line is receiver surface at depth $z_r=1000$ m. Point source location for data
    generation at $(3500, 3500)$ m indicated with star. Overlain with rays from point
    source to receiver surface.}
  \label{fig:olensrays0}
\end{figure}

@@

Figure \ref{fig:dsrcphh0} displays an extended (pressure) source $h_s$, in the form of
traces in the spatial range 2000 m $ \le x_r \le $ 6000 m and time
range 0 s $\le t \le $ 2 s. The modeling operator output $Sh_s$,
obtained from the causal solution $(p, \bv)$ of \ref{eqn:awepm}
with this choice of $h_s$ and $f_s=0,$ by application of the sampling
operator $P_r$ to $p$, is shown in Figure \ref{fig:dfwdplh0}.

@@

\begin{figure}
  \centering
  \subfloat[\label{fig:drecplh0}]{\includegraphics[width=0.32\textwidth]{drecplh0.pdf}}
  \subfloat[\label{fig:dsrcphh0}]{\includegraphics[width=0.32\textwidth]{dsrcphh0.pdf}}
  \subfloat[\label{fig:dfwdplh0}]{\includegraphics[width=0.32\textwidth]{dfwdplh0.pdf}}
  \caption{Data generated using configuration in Figure
    \ref{fig:olensrays0}. (a) traces from point source at $(3500,3500)$
    with $[1, 2, 7.5, 12.5]$ Hz zero phase trapezoidal bandpass
    filter wavelet, delayed by $0.5$ s. (b) Equivalent extended source on
    the source surface at depth $z_s=3000$ m. (c) Gather generated by
    equivalent source shown in (b). Color scale is same for (a) and (c).}
\end{figure}

>>>>

\documentclass[12pt]{geophysics}

@@

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{alg}{Algorithm}
\newtheorem{remark}{Remark}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}

@@

\maketitle
\begin{abstract}
Source extension is a reformulation of inverse problems in wave
propagation, that at least in some cases leads to computationally
tractable iterative solution methods. The core subproblem in all
source extension methods is the solution of a linear inverse problem
for a source (right hand side in a system of wave equations) through
minimization of data error in the least squares sense with soft
imposition of physical constraints on the source via an additive
quadratic penalty. For an acoustic formulation for sources supported
on a surface, with a soft contraint enforcing concentration at a
point, a variant of the time reversal method from photoacoustic
tomography provides an approximate solution. This approximate inverse
can be used to precondition Krylov space iteration for rapid
convergence to the solution of the core subproblem in this
setting. Numerical examples illustrate the effectiveness of this preconditioner.
\end{abstract}

@@

The modeling operator $PL[{\bf c}]^{-1}$ and the penalty operator $A$ defined in \ref{eqn:penop} are linear, so the source
subproblem is a linear least squares problem. Under some additional
assumptions to be described below, I shall show how to construct an
accurate approximate solution operator for problem
\ref{eqn:esis}. This approximate solution operator may be used to
accelerate (``precondition'') Krylov space methods for the solution of the surface source
subproblem \ref{eqn:esis}. I will fully describe a preconditioner for a special
case of the source subproblem \ref{eqn:esis}, in which ${\bf u}$ is an
acoustic field, $L[{\bf c}]$ is the wave operator of linear
acoustodynamics. Numerical examples in this setting suggest the
effectiveness of this acceleration.

@@

A simplified mathematical
translation of this medical imaging task is to infer the initial
excess pressure distribution over a fluid-containing region at time
$t=0$ from measurements of the pressure on a surface enclosing the
fluid over a time interval $0 \le t \le t_{\rm max}$. The time reversal method presumes that the
pressure field has returned to equilibrium (zero excess pressure), or
close to it, in the fluid-containing region at the
final time $t=t_{\rm max}$ Then the field can be (at least
approximately) viewed as the solution of the backwards-in-time initial boundary value
problem with zero final conditions at $t=t_{\rm max}$, and boundary
values given by the measurements. Evolving the field backwards in time
to $t=0$ thus solves the problem. Except in special
circumstances, the pressure field never actually vanishes at finite
time, so the solution is approximate.

@@

First, reverse-time propagation can be localized via ray theory,
within high-frequency asymptotic approximation. This step requires
some assumptions about the ray fields: all rays responsible for
significant energy in the receiver data must arrive from the source
surface, and must cross the source and receiver surfaces
transversally, that is, making non-zero angles. The surfaces must be
oriented, with a choice of (``outward'') unit normal vector. The
energetic rays must emanate
from the source surface inwards (``incoming'' rays), and
arrive at the receiver surface outwards (``outgoing'' rays). Then both surfaces
can be treated locally as
boundaries of propagation domains. 
With these assumptions, reverse-time propagation of the receiver data
closely approximates the acoustic fields near the source
surface. Next, I show that for a causal field with a surface pressure source (with a
continuous pressure field across the surface), the source amplitude is
proportional to the jump in the (particle) velocity field. Moreover,
to leading order in frequency, the velocity field switches sign at the
surface, so the jump is just twice the sampled value on the
surface. Thus constructing a source from the velocity trace on the
receiver surface, reverse-time propagation of this source, and reading off the velocity
field on the source surface yields a source that reproduces the
pressure data, within an asymptotically small error.

This is the essence of the approximate solution of the problem
\ref{eqn:esis} for penalty weight $\alpha=0$. To see how Krylov
iteration might be accelerated, I note that reverse-time propagation of
the receiver surface source to pressure on the source surface is the transpose (adjoint) of the
forward-time source-to-pressure propagation - this is a very simple
version of the adjoint state construction \cite[]{Plessix:06}. So the
approximate inverse may be described as: conversion of pressure to source at the
receiver surface, followed by application of the modeling operator adjoint,
followed by the conversion of pressure to source at the source
surface. This sequence precisely describes the adjoint of the modeling
operator with respect to weighted norms in the spaces of source and
receiver data, with the weight operators being the
pressure-to-source operators on the two surfaces. That this construction results in an
approximate inverse means that the modeling operator is approximately
unitary with respect to these weighted norms, which in turn implies
that a properly constructed preconditioned conjugate gradient algorithm should converge
rapidly.

Finally, the $\alpha=0$ case is not sufficient: the penalty operator
$A$ is an essential component of the nonlinear inverse propblem
\ref{eqn:esi}. As it turns out, $A$ commutes with the other operators
involved - approximately, but that is enough. Therefore the effect of
$A$ can be compensated with an easily-computed factor, whence the
preconditioned acceleration extended to the case $\alpha > 0$.

This pressure-to-source map is closely related to the ``hyperbolic
Dirichlet-to-Neumann'' operator that plays a prominent role in
photoacoustic tomography and other wave inverse problems
\cite[]{Rachele:00,StefUhl:05}. In this terminology, ``Dirichlet''
refers to pressure restricted to a surface, and ``Neumann'' to the
normal derivative of pressure on that surface. Since the latter is
proportional to normal particle acceleration thanks the equations of motion,
and normal particle velocity is proportion to source, the relation to the
pressure-to-source map as described in this paper is clear. A number
of authors have described preconditioners for Least Squares Migration
with subsurface offset extension in which the Dirichlet-to-Neumann or
pressure-to-source map plays a central role
\cite[]{tenKroode:12,HouSymes:15,Herve2017}. It also turns up in hidden
form in the work of Yu Zhang and collaborators on ``true amplitude''
migration
\cite[]{YuZhang:14,TangXuZhang:13,XuWang:2012,XuZhangTang:11,Zhang:SEG09}.

The role of the pressure-to-source map in these constructions is quite
similar to the role described here. Reverse Time Migration, in its
basic form, broadcasts pressure traces into the propagation region,
backwards in time. Where this backward propagated energy intersects
the forward propagated Green's function, an imaging condition outputs
an image amplitude map, which is not however an approximate inversion,
as is widely understood, except in a crude kinematic sense. The
various ``true amplitude'' or approximate least squares migrations,
apply the pressure-to-source operator (either explicitly, or
implicitly via switching boundary conditions) {\em before}
reverse-time broadcasting. Since the data are presumed to be traces of
the forward scattered (Born) pressure field, the backward propagated
wavefield must approximate the forward scattered field,
time-reversed. This reconstructed scattered field combines at
scattering points with the Green's function to yield an approximate
inversion for the materal parameter perturbations. Moreover, this
approximate Born inversion is also adjoint to the Born scattering operator in weighted norms, one
of the weight operators being the pressure-to-source map, hence an
effective preconditioner for Krylov space methods for linear least
squares inversion, that is, Least Squares Migration
\cite[]{HouSymes:16}. Thus the present construction bears a
considerable resemblance to these ``true-amplitude'' reverse-time
migration methods.

The discussion in this paper is formal and incomplete, in the sense
that some important mathematical underpinnings are only
sketched. The transport of high-frequency energy along rays is a
crucial ingredient in the constructions explained here. Of course,
such transport only occurs in material parameter models that vary
slowiy in space on the wavelength scale, and even then only
approximately. Rather than provide explicit
smoothness constraints and asymptotic error estimates, I will use
these concepts in an intuitive way, using ``approximate'' and
``asymptotic'' and the symbol ``$\approx$'' to suggest asymptotically
negligible errors not explicitly assessed, and refer to such errors as
``negligible''.

I will treat the modeling operator $PL[{\bf c}]^{-1}$ as if
it mapped square integrable surface sources to square integrable sampled
data. This is not true in full generality: while the surface source
problem has distribution solutions, they are not generally square
integrable (finite acoustic field energy). Even if the solutions have
finite energy, they do not in general have well-defined
restrictions to lower-dimensional sets. In other words, the action of
the sampling
operator $P$ on the receiver surface is not well-defined for arbitrary
finite-energy acoustic fields. Thus the modeling operator envisioned above
is not well-defined, strictly speaking.
This phenomenon is related to the ill-posedness of wave equations as
evolution equations in spatial variables, an observation attributed to
Hadamard (see \cite{CourHil:62}, Chapter 6, section 17). A number of
authors have described precise forms of the ray conditions described
above, and shown how these conditions lead to the desired behaviour of
the modeling operator, that is, mapping of finite energy sources to
finite energy data, or equivalent properties
\cite[]{Payn:75,Symes:83,Lasi:86,LasLionsTrig:86,Lasi:87, BaoSy:91b}.
Elaboration of these mathematical details is beyond the scope of this
paper, which aims instead to explore the algorithmic consequences of
the mathematical structure implied by the incoming/outgoing ray constrants..

@@

Most of what follows is valid for any space dimension $n >0$. I will
describe the theory for $n=3$, write $\bx=(x,y,z)^T$ for the spatial
coordinate vector, and similarly for the 
particle velocity $\bv = (v_x,v_y,v_z)^T$. For computational convenience, the
examples are two-dimensional.

@@

For
acoustic modeling, surface sources are combinations of constitutive law
defects and loads normal to the surface, localized on $\Sigma_s$. That
is, right-hand sides in the system $L\bu=\bff$ take the form
$\bff(\bx,t) = (h_s\delta_{\Sigma_s},
f_s{\bf n}_s\delta_{\Sigma_s})^T$ for scalar defect $h_s$ and normal
load $f_s$. With the choice $L$ given in
\ref{eqn:awemat}, the causal system $L\bu=\bff$
takes the form
\begin{eqnarray}
\label{eqn:awepm}
\frac{1}{\kappa}\frac{\partial p}{\partial t} & = & - \nabla \cdot \bv +
h_s \delta_{\Sigma_s}, \nonumber \\
\rho\frac{\partial \bv}{\partial t} & = & - \nabla p +
                                                f_s{\bf n}_s \delta_{\Sigma_s},\nonumber \\
p & =& 0 \mbox{ for }  t \ll 0,\nonumber\\ 
\bv & = & 0 \mbox{ for } t \ll 0.
\end{eqnarray}
in which $\delta_{\Sigma_s}$ is the delta function on $\Sigma_s$, and
similarly for $\Sigma_r$.

@@

Extended forward modeling consists in solving \ref{eqn:awepm} and
sampling the solution components at a receiver surface $\Sigma_r$ with
(outward) unit normal field ${\bf n}_r$.
$P_s$ and $P_r$ are the sampling operators on $\Sigma_s$ and
$\Sigma_r$ respectively. In
practice, sampling necessarily occurs at a discrete array of points
(trace locations). In this
theoretical discussion, I will neglect the finite sample rate, and
regard the data, for example $P_rp$, as continuously sampled. The
output samples are necessarily muted, that is, non-zero only over a space-time domain
of finite extent. This mute, and any tapering applied to the data
traces, are regarded as part of the sampling operators $P_s,P_r$.

@@

Denote by $\Pi_i, i=0,1$ the projection on the first,respectively
second, component of a vector in $\bR^2$:
\begin{equation}
  \label{eqn:pidef}
  \Pi_0 (a,b)^T = a, \, \Pi_1(a,b)^T= b.
\end{equation}
The forward modeling
operator from pressure source to pressure trace is
\begin{equation}
  \label{eqn:sdef}
  S = \Pi_0 {\cal S}\Pi_0^T 
\end{equation}
%and the forward modeling operator from velocity source (normal force)
%to velocity trace is
%\begin{equation}
%  \label{eqn:vdef}
%  V_{z_s,z_r} = \Pi_1 {\cal S}_{z_s,z_r} \Pi_1^T 
%\end{equation}

@@

The two examples mentioned in the introduction illustrate the setting
just described. The first example (Figure \ref{fig:olensrays0}) embeds
an acoustic (low-velocity) lens between a source surface
$\Sigma_s= \{z = z_s=3000 \mbox{ m}\}$ (orange horizontal line) m and
receiver surface $\Sigma_r= \{z=z_r=1000 \mbox{ m}\}$ (yellow
horizontal line).  The data to be used in this example results from a
point source at $(3500, 3500)$ m (orange star). I have overlain
the rays connecting this source point with the data portion of the
receiver surface: evidently all of the rays involved cross the source
and receiver surfaces transversally. The outward normal vectors are chosen as
${\bf n}_s={\bf e}_z=(0,1)^T$ on the source surface and ${\bf
  n}_r=-{\bf e}_z$ on the receiver surface.
One sees from the Figure that the ray field carrying high frequency
energy is incoming at the source surface,
outgoing at the receiver surface.

\begin{figure}
  \includegraphics[width=\textwidth]{olensrays0.pdf}
  \caption{Bulk modulus with acoustic lens. Color scale unit is
    GPa. Orange horizontal line is source surface at depth $z_s= 3000$
    m, yellow horizontal
    line is receiver surface at depth $z_r=1000$ m. Point source location for data
    generation at $(3500, 3500)$ m indicated with star. Overlain with rays from point
    source to receiver surface. Outward normal points down at source
    surface, up at receiver surface, so rays are incoming at source
    surface, outgoing at receiver surface.}
  \label{fig:olensrays0}
\end{figure}

@@

Figure \ref{fig:dhshh0} displays an extended (pressure) source $h_s$, in the form of
traces in the spatial range 2000 m $ \le x_r \le $ 6000 m and time
range 0 s $\le t \le $ 2 s. The modeling operator output $Sh_s$,
obtained from the causal solution $(p, \bv)$ of \ref{eqn:awepm}
with this choice of $h_s$ and $f_s=0,$ by application of the sampling
operator $P_r$ to $p$, is shown in Figure \ref{fig:dfwdplh0}.

@@

\begin{figure}
  \centering
  \subfloat[\label{fig:drecplh0}]{\includegraphics[width=0.32\textwidth]{drecplh0.pdf}}
  \subfloat[\label{fig:dhshh0}]{\includegraphics[width=0.32\textwidth]{dhshh0.pdf}}
  \subfloat[\label{fig:dfwdplh0}]{\includegraphics[width=0.32\textwidth]{dfwdplh0.pdf}}
  \caption{Data generated using configuration in Figure
    \ref{fig:olensrays0}. (a) traces from point source at $(3500,3500)$
    with $[1, 2, 7.5, 12.5]$ Hz zero phase trapezoidal bandpass
    filter wavelet, delayed by $0.5$ s. (b) Equivalent extended source on
    the source surface at depth $z_s=3000$ m. (c) Gather generated by
    equivalent source shown in (b). Color scale is same for (a) and (c).}
\end{figure}
